Untitled

### 2024-10-04 code/manuscript_altanalysis.Rmd
closures too!

```{r}
library(readxl)
library(did2s)
library(tidyverse)
library(directlabels)
library(gtsummary)
library(kableExtra)
library(ggpubr)
library(tigris)
library(usmap)
library(ivreg)
library(fixest)
library(sandwich)
library(lmtest)
```

@@

dfc <- dfc %>% 
  mutate(Qcomp = (dm.rate + htn.rate + crc.rate + pap.rate + childbmi.rate + adultbmi.rate ) / 6 * 100, 
         Qcomp.w = (dm.rate * dm.total + htn.rate * htn.total + crc.rate * crc.denom + pap.rate * pap.denom + childbmi.rate * childbmi.denom + adultbmi.rate * adultbmi.denom ) / 
           ( dm.total +  htn.total +  crc.denom +  pap.denom +  childbmi.denom +  adultbmi.denom ) * 100, 
               CostPerPat = ngla_expense / Total, 
               months_of_cash = as.numeric(months_of_cash),
               ngla_prog_serv_revenue = as.numeric(ngla_prog_serv_revenue),
               quick_ratio = as.numeric(quick_ratio),
               Liquidity = (Assets - Liabilities) / Assets,
               Profitability = ngla_net_gain_loss / ngla_expense * 100, #Assets, 
               Efficiency = ngla_revenue / Assets, #ngla_revenue / Assets,
               NetWorth = (Assets - Liabilities) / Liabilities, 
            Pexpenses = ngla_prog_serv_revenue / ngla_prog_expense, 
            LiabPctAssets = Liabilities / Assets * 100, 
            Eff = ngla_prog_serv_revenue / Assets, 
            LiabPctRev = ngla_prog_serv_revenue / ngla_expense)

@@

df.temptemp <- df.mt %>% 
  rename(ID = ID.x) %>% 
  mutate(Closed = ifelse(lead(is.na(Sites)), 1, 0)) %>% 
  select(ClinicHHI, Att, `BHCMIS ID`, value, ID, MS, Year, Pct.NHA:Pct.Homeless, Qcomp, Total, Altman, 
         htn.rate, dm.rate, pap.rate, crc.rate, childbmi.rate, adultbmi.rate, Complexity, Intensity, Pct.MCD, Pct.MCR, Pct.COM, Pct.None, CostPerPat, ngla_revenue, ngla_expense, Profitability, LiabPctAssets, Staff) %>% 
  unique()

@@

  filter(value_min == value_max | value_last > value_first) %>% 
  mutate(value = ifelse(value >= 6, 6, value))

@@

for (k in 1:21){

@@

core.iv.estimates <- data.frame(cbind(colnames(df.temptemp)[17:37], (coeffs[-1,]))) %>% 
  #select(V1, Estimate, Pr...t.., Estimate.1, Pr...t...1) %>% 
  mutate(Estimate = round(as.numeric(Estimate), 2), `95% CI` = paste(round(as.numeric(Estimate) - 1.96 * as.numeric(`Std..Error`), 2), ", ", round(as.numeric(Estimate) + 1.96 * as.numeric(`Std..Error`), 2), sep = "")) %>% 
  select(V1, Estimate, `95% CI`)

@@

core.iv.estimates %>% 
  kable("html", booktabs = T, col.names = c("Outcome","Estimate","95% CI")) %>% 
  kable_styling(latex_options = c("striped")) %>%
  pack_rows(index = c("Overall Impacts" = 3, "Quality of Care" = 6, "Patient Complexity" = 2, "Payer Mix" = 4, "Finances" = 6)) %>% 
  add_header_above(c("IV-Estimated LATE Effects of Competitive Shocks on Incumbent Clinics" = 3))

@@

## First Stage DID Estimates

@@

## First Stage DID Event Study
```{r}

 event_study(data = df.temptemp ,
  yname = "ClinicHHI", idname = "ID", tname = "Year", gname = "Att", estimator = "all", xformla = ~Pct.NHA + Pct.NHB + Pct.HL + Pct.LessThan100pctFPL + Pct.FarmWorker + Pct.Homeless) %>%
  filter(abs(term) <= 5) %>%
  #filter(estimator == "Gardner (2021)") %>%
  ggplot(aes(x = term, y = estimate, color = estimator)) + geom_point(position = position_dodge(0.9), size = 3) + geom_errorbar(aes(ymin=estimate-1.96*std.error, ymax=estimate+1.96*std.error), width=.25, position = position_dodge(0.9), size = 0.8)  +
  #facet_wrap(~V1, scales = "free_y", ncol = 3) +
  theme_tufte_revised() + geom_vline(xintercept=-0.5, color="blue", linetype="dashed", alpha=.6)+ #theme(legend.position = "none", plot.title = element_text(size=18, face='bold')) +
  geom_hline(yintercept=0, color="black", linetype="dashed", alpha=.5) + ggtitle("FirstStage: ") + xlab("Year Relative to Competitive Shock") + ylab("Percentage Points")

@@

df.temptemptemp <- df.temptemp %>% select(ClinicHHI, Qcomp, MS, value, Year, ID, Pct.NHA:Pct.COM) %>% ungroup() %>% filter(complete.cases(.))

@@

@@

```{r}
df.temptemp %>% #left_join(df.temptemp %>% filter(Tx == 1) %>% summarise(Att = min(Year)), by = c("ID")) %>% 
               mutate(Att = ifelse(is.na(Att), 0, Att)) %>% 
  mutate(value = ifelse(value >= 6, 6, value)) %>% 
  group_by(value, Year) %>% 
  summarise(MS = mean(MS, na.rm = TRUE)) %>% 
  ggplot(aes(x = Year, y = MS, group = value, color = as.factor(value))) + geom_point() + geom_line()
```

@@

df.z

@@

df.z.temp <- df.z %>% mutate(Tx = ifelse(value - 1 > 1, 1, value - 1)) %>% filter(!is.na(Tx)) %>% mutate(DirectlyTreated = ifelse(Att.z > 2010, 1, 0)) %>% 
        mutate(Pct.MCD = Medicaid / NumberofPatients * 100, 
               Pct.MCR = Medicare / NumberofPatients * 100, 
               Pct.COM = Private / NumberofPatients * 100,
               Pct.None = Uninsured / NumberofPatients * 100 ) %>% 
  select(NumberofPatients, Pct.MCD:Pct.None, Total.Hours, Sites, Mean.Hours, Perminance, Schedule, Calendar, Tx, DirectlyTreated, ID, Year)

@@

  df.z.temp2 <- df.z.temp
  colnames(df.z.temp2)[k] <- "Outcome"
  model <-  did2s(
      data = df.z.temp2, 
      yname = "Outcome", 
      treatment = "Tx",
        first_stage = ~ 1 | ID + Year , #Pct.NHA + Pct.NHB + Pct.HL + Pct.LessThan100pctFPL + Pct.MCD + Pct.MCR + Pct.None | ID + Year , 
        second_stage = ~ i(Tx, ref = FALSE)  + I(Tx * DirectlyTreated),
        cluster_var = "ID",
      verbose = TRUE
    )

@@

@@

>>>>

```{r}
library(readxl)
library(did2s)
library(did)
library(tidyverse)
library(directlabels)
library(gtsummary)
library(kableExtra)
library(ggpubr)
library(tigris)
library(usmap)
library(ivreg)
library(fixest)
library(sandwich)
library(lmtest)
```

@@

dfc <- dfc %>% 
  mutate(Qcomp = (dm.rate + htn.rate + crc.rate + pap.rate + childbmi.rate + adultbmi.rate ) / 6 * 100, 
         Qcomp.w = (dm.rate * dm.total + htn.rate * htn.total + crc.rate * crc.denom + pap.rate * pap.denom + childbmi.rate * childbmi.denom + adultbmi.rate * adultbmi.denom ) / 
           ( dm.total +  htn.total +  crc.denom +  pap.denom +  childbmi.denom +  adultbmi.denom ) * 100, 
               CostPerPat = ngla_expense / Total, 
               months_of_cash = as.numeric(months_of_cash),
               ngla_prog_serv_revenue = as.numeric(ngla_prog_serv_revenue),
               quick_ratio = as.numeric(quick_ratio),
               Liquidity = (Assets - Liabilities) / Assets,
               Profitability = ngla_revenue / ngla_expense * 100, #Assets, 
               Efficiency = ngla_revenue / Assets, #ngla_revenue / Assets,
               NetWorth = (Assets - Liabilities) / Liabilities, 
            Pexpenses = ngla_prog_serv_revenue / ngla_prog_expense, 
            LiabPctAssets = Liabilities / Assets * 100, 
            Eff = ngla_prog_serv_revenue / Assets, 
            LiabPctRev = ngla_prog_serv_revenue / ngla_expense)

@@

df.temptemp <- df.mt %>% 
  rename(ID = ID.x) %>% 
  mutate(Closed = ifelse(lead(is.na(Sites)), 1, 0)) %>% 
  select(ClinicHHI, Att, `BHCMIS ID`, value, ID, MS, Year, Pct.NHA:Pct.Homeless, Qcomp, Total, Altman, logPatsPerStaff, 
         htn.rate, dm.rate, pap.rate, crc.rate, childbmi.rate, adultbmi.rate, Complexity, Intensity, Pct.MCD, Pct.MCR, Pct.COM, Pct.None, CostPerPat, quick_ratio, months_of_cash, Profitability, LiabPctAssets, Staff, Closed) %>% 
  unique()

@@

  filter(value_min == value_max | value_last > value_first) %>% 
  mutate(value = ifelse(value >= 6, 6, value)) #%>% filter(Year < 2020)

@@

for (k in 1:22){

@@

core.iv.estimates <- data.frame(cbind(colnames(df.temptemp)[17:38], (coeffs[-1,]))) %>% 
  #select(V1, Estimate, Pr...t.., Estimate.1, Pr...t...1) %>% 
  mutate(Estimate = round(as.numeric(Estimate), 2), `95% CI` = paste(round(as.numeric(Estimate) - 1.96 * as.numeric(`Std..Error`), 2), ", ", round(as.numeric(Estimate) + 1.96 * as.numeric(`Std..Error`), 2), sep = "")) %>% 
  select(V1, Estimate, `95% CI`)

@@

core.iv.estimates %>% 
  kable("html", booktabs = T, col.names = c("Outcome","Estimate","95% CI")) %>% 
  kable_styling(latex_options = c("striped")) %>%
  pack_rows(index = c("Overall Impacts" = 4, "Quality of Care" = 6, "Patient Complexity" = 2, "Payer Mix" = 4, "Finances" = 6)) %>% 
  add_header_above(c("IV-Estimated LATE Effects of Competitive Shocks on Incumbent Clinics" = 3))

feols(I(Closed * 100) ~  Pct.NHB + Pct.HL + Pct.LessThan100pctFPL + Pct.FarmWorker + Pct.Homeless | I((10000 - ClinicHHI)/1000) ~ Tx + factor(Year) + factor(ID) , cluster = "ID", data = df.temptemp)

@@

## First Stage DID Estimates and Event Study

@@

 event_study(data = df.temptemp ,
  yname = "ClinicHHI", idname = "ID", tname = "Year", gname = "Att", estimator = "all", xformla = ~Pct.NHA + Pct.NHB + Pct.HL + Pct.LessThan100pctFPL + Pct.FarmWorker + Pct.Homeless) %>%
  filter(abs(term) <= 6) %>%
  #filter(estimator == "Gardner (2021)") %>%
  filter(estimator == "Callaway and Sant'Anna (2020)") %>%
  ggplot(aes(x = term, y = estimate, color = estimator)) + geom_point(position = position_dodge(0.9), size = 3) + geom_errorbar(aes(ymin=estimate-1.96*std.error, ymax=estimate+1.96*std.error), width=.25, position = position_dodge(0.9), size = 0.8)  +
  #facet_wrap(~V1, scales = "free_y", ncol = 3) +
  theme_tufte_revised() + geom_vline(xintercept=-0.5, color="blue", linetype="dashed", alpha=.6)+ #theme(legend.position = "none", plot.title = element_text(size=18, face='bold')) +
  geom_hline(yintercept=0, color="black", linetype="dashed", alpha=.5) + ggtitle("FirstStage: ") + xlab("Year Relative to Competitive Shock") + ylab("Percentage Points")

@@

## TWFE First Stage
df.temptemptemp <- df.temptemp %>% select(ClinicHHI, Qcomp, MS, value, Year, ID, Pct.NHA:Pct.None) %>% ungroup() %>% filter(complete.cases(.))

@@

# ## Callaway-Sant'Anna First Stage
# 
# example_attgt <- att_gt(yname = "ClinicHHI",
#                         tname = "Year",
#                         idname = "ID",
#                         gname = "Att",
#                         xformla = ~Pct.NHB + Pct.HL + Pct.LessThan100pctFPL + Pct.FarmWorker + Pct.Homeless,
#                         data = df.temptemp, 
#                         clustervars = "ID"
#                         )
# 
# # summarize the results
# aggte(example_attgt)

@@

```{r}
df.temptemp %>% #left_join(df.temptemp %>% filter(Tx == 1) %>% summarise(Att = min(Year)), by = c("ID")) %>% 
               mutate(Att = ifelse(is.na(Att), 0, Att)) %>% 
  mutate(value = ifelse(value >= 6, 6, value)) %>% 
  group_by(value, Year) %>% 
  summarise(HHI = mean(ClinicHHI, na.rm = TRUE)) %>% 
  ggplot(aes(x = Year, y = HHI, group = value, color = as.factor(value))) + geom_point() + geom_line()
```

@@

@@

df.z.temp <- df.z %>% mutate(Tx = ifelse(value - 1 > 1, 1, value - 1)) %>% filter(!is.na(Tx)) %>% mutate(DirectlyTreated = ifelse(Att.z > 2010, 1, 0)) %>% 
  filter(!(Att == 0 & DirectlyTreated == 1)) %>% 
        mutate(Pct.MCD = Medicaid / NumberofPatients * 100, 
               Pct.MCR = Medicare / NumberofPatients * 100, 
               Pct.COM = Private / NumberofPatients * 100,
               Pct.None = Uninsured / NumberofPatients * 100 ) %>% 
  select(NumberofPatients, Pct.MCD:Pct.None, Total.Hours, Sites, Mean.Hours, Perminance, Schedule, Calendar, Tx, DirectlyTreated, ID, Year, Zip, Att)

@@

  df.z.temp2 <- df.z.temp #%>% filter(Att == 0 | Att > 2010)
  colnames(df.z.temp2)[k] <- "Outcome"
  model <-  did2s(
      data = df.z.temp2, 
      yname = "Outcome", 
      treatment = "Tx",
        first_stage = ~ 1 | ID + Year , #Pct.NHA + Pct.NHB + Pct.HL + Pct.LessThan100pctFPL + Pct.MCD + Pct.MCR + Pct.None | ID + Year , 
        second_stage = ~ i(Tx * (1 - DirectlyTreated), ref = FALSE)  + I(Tx * DirectlyTreated),
        cluster_var = "ID",
      verbose = TRUE
    )

@@

## WATCH OUT FOR THE INTERACTION INTERPRETATION

@@

## Event Study

```{r}
 event_study(data = df.z.temp %>% filter(Att == 0 | DirectlyTreated == 0),
  yname = "Total.Hours", idname = "ID", tname = "Year", gname = "Att", estimator = "all", xformla = ~1) %>%
  filter(abs(term) <= 6) %>%
  filter(estimator == "Gardner (2021)") %>%
  #filter(estimator == "Callaway and Sant'Anna (2020)") %>%
  ggplot(aes(x = term, y = estimate, color = estimator)) + geom_point(position = position_dodge(0.9), size = 3) + geom_errorbar(aes(ymin=estimate-1.96*std.error, ymax=estimate+1.96*std.error), width=.25, position = position_dodge(0.9), size = 0.8)  +
  #facet_wrap(~V1, scales = "free_y", ncol = 3) +
  theme_tufte_revised() + geom_vline(xintercept=-0.5, color="blue", linetype="dashed", alpha=.6)+ #theme(legend.position = "none", plot.title = element_text(size=18, face='bold')) +
  geom_hline(yintercept=0, color="black", linetype="dashed", alpha=.5) + xlab("Year Relative to Competitive Shock") + ylab("Percentage Points")

 event_study(data = df.z.temp %>% filter(Att == 0 | DirectlyTreated == 1),
  yname = "Total.Hours", idname = "ID", tname = "Year", gname = "Att", estimator = "all", xformla = ~1) %>%
  filter(abs(term) <= 6) %>%
  filter(estimator == "Gardner (2021)") %>%
  #filter(estimator == "Callaway and Sant'Anna (2020)") %>%
  ggplot(aes(x = term, y = estimate, color = estimator)) + geom_point(position = position_dodge(0.9), size = 3) + geom_errorbar(aes(ymin=estimate-1.96*std.error, ymax=estimate+1.96*std.error), width=.25, position = position_dodge(0.9), size = 0.8)  +
  #facet_wrap(~V1, scales = "free_y", ncol = 3) +
  theme_tufte_revised() + geom_vline(xintercept=-0.5, color="blue", linetype="dashed", alpha=.6)+ #theme(legend.position = "none", plot.title = element_text(size=18, face='bold')) +
  geom_hline(yintercept=0, color="black", linetype="dashed", alpha=.5) + xlab("Year Relative to Competitive Shock") + ylab("Percentage Points")

## Are the closer and farther areas different?
```{r}

df.z.temp %>% 
  #select(Zip, Year, DirectlyTreated, Att) %>% 
  mutate(Status = ifelse(Att == 0, NA, 
                         ifelse(Att > 0 & DirectlyTreated == 0, "IndirectlyTreated","DirectlyTreated"))) %>% 
  inner_join(census, by = c("Zip" = "ZipCode", "Year" = "YEAR")) %>% 
  filter(Year == 2012) %>% 
  mutate(PctMedicaid = Medicaid / MCD.denom * 100, 
         PctUninsured = Uninsured / Unins.denom * 100, 
         PctPoverty = IncomeBelowPoverty / PovertyStatusDet * 100, 
         PctBlack = Black / TotalPop * 100, 
         PctHispanic = Hispanic / TotalPop * 100) %>% 
    select(Status, TotalPop, PctMedicaid:PctHispanic) %>% 
    tbl_summary(by = Status, 
                statistic = list(all_continuous() ~ "{mean} ({sd})"), missing = "no") %>% 
  add_p()
